---
title: "p8105_hw6_rd3096"
author: Timothy Du
output: github_document
---

# load necessary pakages 
```{r setup}
library(tidyverse)
library(rnoaa)
library(ggplot2)
library(modelr)
library(mgcv)
library(broom)
library(purrr)


set.seed(1)


knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## problem 1

we need to download the dataset using the code chunk below
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

str(weather_df)

weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point() 
```

Next, we will perform bootstrap sampling and regression analysis.

First, we need to write a function to create the two quantities that we are interested.
```{r}
compute_metrics = function(weather_sample) {
  model = lm(tmax ~ tmin, data = weather_sample)
  r_squared = glance(model)$r.squared
  coefficients = tidy(model)
  intercept = coefficients %>% 
    filter(term == "(Intercept)") %>% 
    pull(estimate)
  
  slope = coefficients %>%
    filter(term == "tmin") %>% 
    pull(estimate)
  
  log_beta_product = log(intercept * slope)
  
  tibble(
    r_squared = r_squared,
    log_beta_product = log_beta_product
  )
}
```

we will perform bootstramp using the map function
```{r}
bootstrap_results = 
  weather_df %>%
  modelr::bootstrap (n = 5000) %>% 
  mutate(
    strap = map(strap, as_tibble),
    results = map(strap, compute_metrics)
    ) %>%
  unnest(results)
```

Final step, we will summarize the results and visualize using ggplot
```{r}
bootstrap_summary =
  bootstrap_results %>%
  summarise(
    mean_r_squared = mean(r_squared),
    sd_r_squared = sd(r_squared),
    mean_log_beta_product = mean(log_beta_product),
    sd_log_beta_product = sd(log_beta_product)
  )


bootstrap_results %>%
  pivot_longer(
    cols = c(r_squared, log_beta_product), 
    names_to = "metric", 
    values_to = "value"
    ) %>%
  ggplot(aes(x = value)) +
  geom_density(fill = "blue", alpha = 0.7) +
  facet_wrap(. ~ metric, scales = "free") +
  theme_minimal() +
  labs(
    title = "Bootstrap Distributions of r^2 and log(beta0 * beta1)",
    x = "Value",
    y = "Frequency"
  )
```
1. Distribution of log(beta_0*beta_1)

* The plot on the left represents the bootstrap distribution of this log_product, where beta_0 is the intercept and beta_1 is the slope.

* Shape: The distribution is approximately symmetric and bell-shaped, indicating that the log-transformed product of the intercept and slope is fairly stable across the bootstrap samples.

* Range: The values are centered around 2.0, with a slight spread from approximately 1.95 to 2.10.

* Interpretation: The logarithm of the product of the intercept and slope is consistent and suggests a central tendency (average value) close to 2.0.

2. Distribution of r_square

* The plot on the right represents the bootstrap distribution of r_square, the proportion of variance in tmax explained by tmin.

* Shape: This distribution is also symmetric and bell-shaped, showing that the r_square values are consistently estimated across bootstrap samples.

* Range: The values are tightly centered around 0.91, with a slight spread from approximately 0.88 to 0.94.

* Interpretation: The r_square values indicate a strong linear relationship between tmin and tmax, with most bootstrap samples confirming a high proportion of variance explained.

we will construct the 95% CI

```{r}
bootstrap_conf_intervals =
  bootstrap_results %>%
  summarise(
    r_squared_lower = quantile(r_squared, 0.025),
    r_squared_upper = quantile(r_squared, 0.975),
    log_beta_product_lower = quantile(log_beta_product, 0.025),
    log_beta_product_upper = quantile(log_beta_product, 0.975)
  )
bootstrap_conf_intervals
```
Based on the output, we know that the 95% confidence interval for r^2 is between 0.894 and 0.927, the  95% confidence interval for log_beta_product is between 1.97 and 2.06.

## problem 2

first, we will read and clean the dataset
```{r}
homicide_data = read_csv("https://raw.githubusercontent.com/washingtonpost/data-homicides/refs/heads/master/homicide-data.csv", na = c("NA","","."))

homicide_clean = 
  homicide_data %>%
  mutate(
    city_state = str_c(city, state, sep = ", "),
         solved = if_else(disposition == "Closed by arrest", 1, 0)) %>% 
  select(city_state,everything()) %>% 
  filter(!city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL")) %>%
  filter(victim_race %in% c("White", "Black")) %>%
  mutate(victim_age = as.numeric(victim_age)) %>% 
  drop_na(victim_age)
```

we will run the glm function for city of Baltimore and tidy the result

```{r}
baltimore_data =
  homicide_clean %>%
  filter(city_state == "Baltimore, MD") %>% 
  select(solved, victim_age, victim_race, victim_sex)


baltimore_glm =
  glm(
  solved ~ victim_age + victim_sex + victim_race, 
  data = baltimore_data, 
  family = binomial
) 
save(baltimore_glm, file = "./results/baltimore_glm.RData")


baltimore_tidy =
  baltimore_glm %>% 
  broom::tidy() %>% 
  knitr::kable(digits=3)
baltimore_tidy
```

we obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing male victims to female victims keeping all other variables fixed.
```{r}
baltimore_or_ci = 
  broom::tidy(
    baltimore_glm, 
    conf.int = TRUE        
  ) %>% 
  mutate(
    OR = exp(estimate),           
    adjusted_conf.low = exp(conf.low), 
    adjusted_conf.high = exp(conf.high)
  ) %>% 
  select(
    term, 
    OR,                          
    adjusted_conf.low,            
    adjusted_conf.high            
  ) %>% 
  knitr::kable(digits = 3)        

baltimore_or_ci
```
1. Interpretation:

Odds Ratio (OR):The adjusted odds ratio (OR) for solving homicides comparing male victims to female victims is 0.426.This means that the odds of solving homicides are about 57.4% lower for male victims compared to female victims, adjusting for age and race.

2. Confidence Interval:

The 95% confidence interval for the OR is [0.324, 0.558]. This indicates that we are 95% confident the true odds ratio lies between 0.324 and 0.558.

Next we will run glm for each of the cities in our dataset
```{r}
city_glm_results =
  homicide_clean %>%
  group_by(city_state) %>%                           
  nest() %>%                                         
  mutate(
    glm_model = map(.x = data, ~ glm(                   
      solved ~ victim_age + victim_sex + victim_race, 
      data = .x, 
      family = binomial)
    ),
    tidy_results = map(glm_model, ~ broom::tidy(    
      .x, 
      conf.int = TRUE, 
      exponentiate = TRUE)
    )
  ) %>%
  unnest(tidy_results) %>%                          
  filter(term == "victim_sexMale") %>%              
  select(city_state, term, estimate, conf.low, conf.high) %>%  
  rename(
    OR = estimate,                                  
    adjusted_conf.low = conf.low,
    adjusted_conf.high = conf.high
  )

city_glm_results 
```

Next, Create a plot that shows the estimated ORs and CIs for each city. Organize cities according to estimated OR, and comment on the plot.

```{r}
city_results = 
  city_glm_results %>% 
  ungroup() %>% 
  mutate(city_state = fct_reorder(city_state, OR)) 

  ggplot(city_results, aes(x = OR, y = city_state)) +
  geom_point(size = 2, color = "blue") +
  geom_errorbarh(aes(xmin = adjusted_conf.low, xmax = adjusted_conf.high), height = 0.2, color = "black") +  
  labs(
    title = "Adjusted Odds Ratios for Solving Homicides (Male vs Female Victims)",
    x = "Adjusted Odds Ratio (OR)",
    y = "City",
    caption = "Error bars represent 95% confidence intervals"
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 10),    
    plot.title = element_text(hjust = 0.5, size = 14) 
  )
```
`r "**_Comments on the Plot_**"`

1. General Trend: Most cities have ORs below 1, indicating male victims are less likely to have their homicides solved compared to female victims. A few cities, like New York, NY, show ORs near or above 1, suggesting no disparity or a slight advantage for male victims.

2. Statistical Significance: Cities where the confidence intervals (CIs) do not include 1 (e.g., Albuquerque, NM) show significant differences. Wide CIs in cities like Baltimore, MD suggest greater uncertainty.

3. Variability: There is noticeable variability across cities, with some (e.g., Stockton, CA) showing clear disparities and others showing near equality or greater uncertainty.

4. Conclusion: The plot highlights disparities across cities, warranting further investigation into the factors influencing these differences.

## problem 3
we will Load the dataset
```{r}
birth_data = read_csv ("./birthweight.csv", na = c("NA","",".")) %>% 
  mutate(
    babysex = factor(babysex, levels = c(1, 2), labels = c("Male", "Female")),
    malform = factor(malform, levels = c(0, 1), labels = c("Absent", "Present")),
    frace = factor(frace, levels = c(1, 2, 3, 4, 8, 9), labels = c("White", "Black", "Asian", "Puerto Rican", "Other", "Unknown")),
    mrace = factor(mrace, levels = c(1, 2, 3, 4, 8), labels = c("White", "Black", "Asian", "Puerto Rican", "Other"))
  )
```

we will fit a linear model
```{r}

birthweight_model = lm(
  bwt ~ gaweeks + ppbmi + wtgain + malform,
  data = birth_data
) 
 
birthweight_model %>% tidy

birthweight_plot_data =
  birth_data %>%
  add_predictions(birthweight_model, var = "fitted_values") %>%
  add_residuals(birthweight_model, var = "residuals") 

  ggplot(birthweight_plot_data, aes(x = fitted_values, y = residuals)) +
  geom_point(alpha = 0.5, color = "black") +  
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") + 
  labs(
    title = "Residuals vs Fitted Values",
    x = "Fitted Values",
    y = "Residuals"
  ) +
  theme_minimal()

```


