---
title: "p8105_hw6_rd3096"
author: Timothy Du
output: github_document
---

# load necessary pakages 
```{r setup}
library(tidyverse)
library(rnoaa)
library(ggplot2)
library(modelr)
library(mgcv)
library(broom)


set.seed(1)


knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## problem 1

we need to download the dataset using the code chunk below
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

str(weather_df)
```

Next, we will perform bootstrap sampling and regression analysis.

First, we need to write a function to create the two quantities that we are interested.
```{r}
compute_metrics = function(weather_sample) {
  model = lm(tmax ~ tmin, data = weather_sample)
  r_squared = glance(model)$r.squared
  coefficients = tidy(model)
  intercept = coefficients %>% 
    filter(term == "(Intercept)") %>% 
    pull(estimate)
  
  slope = coefficients %>%
    filter(term == "tmin") %>% 
    pull(estimate)
  
  log_beta_product = log(intercept * slope)
  
  tibble(
    r_squared = r_squared,
    log_beta_product = log_beta_product
  )
}
```

we will perform bootstramp using the map function
```{r}
bootstrap_results = 
  weather_df %>%
  modelr::bootstrap (n = 5000) %>% 
  mutate(
    strap = map(strap, as_tibble),
    results = map(strap, compute_metrics)
    ) %>%
  unnest(results)
```

Final step, we will summarize the results and visualize using ggplot
```{r}
bootstrap_summary =
  bootstrap_results %>%
  summarise(
    mean_r_squared = mean(r_squared),
    sd_r_squared = sd(r_squared),
    mean_log_beta_product = mean(log_beta_product),
    sd_log_beta_product = sd(log_beta_product)
  )


bootstrap_results %>%
  pivot_longer(
    cols = c(r_squared, log_beta_product), 
    names_to = "metric", 
    values_to = "value"
    ) %>%
  ggplot(aes(x = value)) +
  geom_density(fill = "blue", alpha = 0.7) +
  facet_wrap(. ~ metric, scales = "free") +
  theme_minimal() +
  labs(
    title = "Bootstrap Distributions of r^2 and log(beta0 * beta1)",
    x = "Value",
    y = "Frequency"
  )
```
1. Distribution of log(beta_0*beta_1)

* The plot on the left represents the bootstrap distribution of this log_product, where beta_0 is the intercept and beta_1 is the slope.

* Shape: The distribution is approximately symmetric and bell-shaped, indicating that the log-transformed product of the intercept and slope is fairly stable across the bootstrap samples.

* Range: The values are centered around 2.0, with a slight spread from approximately 1.95 to 2.10.

* Interpretation: The logarithm of the product of the intercept and slope is consistent and suggests a central tendency (average value) close to 2.0.

2. Distribution of r_square

* The plot on the right represents the bootstrap distribution of r_square, the proportion of variance in tmax explained by tmin.

* Shape: This distribution is also symmetric and bell-shaped, showing that the r_square values are consistently estimated across bootstrap samples.

* Range: The values are tightly centered around 0.91, with a slight spread from approximately 0.88 to 0.94.

* Interpretation: The r_square values indicate a strong linear relationship between tmin and tmax, with most bootstrap samples confirming a high proportion of variance explained.

we will construct the 95% CI

```{r}
bootstrap_conf_intervals =
  bootstrap_results %>%
  summarise(
    r_squared_lower = quantile(r_squared, 0.025),
    r_squared_upper = quantile(r_squared, 0.975),
    log_beta_product_lower = quantile(log_beta_product, 0.025),
    log_beta_product_upper = quantile(log_beta_product, 0.975)
  )
bootstrap_conf_intervals
```

